---
title: "ShinyApp - Logistic Function"
author: "JL"
date: "09/2021"
output: slidy_presentation
---
```{r, include = FALSE}

library(dplyr)
## Helper Functions

# LogisticEquation
LogisticEquation <- function(Xn, r){
  
  # Xn, the first position, proportional to
  # the current population 0 < Xn < 1
  
  # r, the growth rate
  
  Xn_1 <- Xn
  Xn_1 <- r * Xn * (1 - Xn)
  Xn_1
}

# Repeated LogisticEquation to return a vector
LogisticLoop <- function(Xi, r, iter = 100){
  
  # Xi, the first position, proportional to
  # the current population 0 < Xi < 1
  
  # r, the growth rate
  
  # iter, number of iterations

  X_vector <- c(Xi)
  Xii <- NULL
  for (index in 1:iter){
    Xii <- LogisticEquation(Xi, r)
    X_vector <- c(X_vector, Xii)
    Xi <- Xii
  }
  X_vector
}

# Calculates Stable solutions for growth rates
StableSolutions <- function(n = 5, iter = 2000, tol = 1e-3){
  # n, power of 2 for bifurcations
  # iter, number of test solution pairs squared
  # tol, tolerance
  
  x <- seq(from = 0, to = 1, length.out = iter)
  
  r1 <- seq(from = 0, to = 3, length.out = iter*0.2)
  r2 <- seq(from = 3, to = 4, length.out = iter*0.8)
  
  r <- c(r1, r2)
  
  solutions <- data.frame(R = c(), X = c())
  
  for (ri in r){
    
    stable <- (ri - 1)/(ri + 1e-16)    # avoid /0
    
    for (xi in x){

      if (abs(xi - stable) <= tol){
        
        stable_pairs <- data.frame(R = c(ri), X = c(xi))
        
        sols <- c(xi)
        sol1_x <- xi
        
        for (index in 1:2^n){
          
          sol2_x <- LogisticEquation(sol1_x, ri)
          
          if ((abs(sol1_x - sol2_x) > tol) & !(sol2_x %in% sols)){
            
            sols <- c(sols, sol2_x)
            stable_pairs <- rbind(stable_pairs, c(ri, sol2_x))
            sol1_x <- sol2_x
            
          } else { 
            
            break
          }
        }
        
        solutions <- rbind(solutions, stable_pairs)
      }
    }
  }
  names(solutions) <- c("R", "X")
  solutions
}

```


## The Logistic Function

The Logistic Function is especially useful in modelling Population Growth.

We can model the logistic function in R like so

```{r}
LogisticEquation <- function(Xn, r){
  
  # Xn, the first position, proportional to
  # the current population 0 < Xn < 1
  
  # r, the growth rate
  
  Xn_1 <- Xn
  Xn_1 <- r * Xn * (1 - Xn)
  Xn_1
}
```
We take an initial population `Xn` and we get the next iteration, `Xn_1`. The growth rate, `r` controls this.

## An example

With an initial population of `0.2` (that is, half of the maximum possible population), and a growth rate of `2` we have the following. It reaches a population cap at about 0.5, after only a few iterations.

```{r echo=FALSE, fig.width = 10}
xVector <- LogisticLoop(0.2, 2, iter = 25)
xMean <- mean(xVector)

plot(xVector, xlab = "Iterations",ylab = "Population",
     col = "red",  ylim = c(0,1))
lines(xVector, col = "red", lwd = 1)
abline(h = xMean, col = "blue")
```

## Growth Rate above 3

If we take the growth rate above `3`, chaotic behaviour begins to happen, regardless of initial population. It reaches a population cap at about 0.5, after only a few iterations. Here is r = 3.5. Notice though that there are repeated solutions...

```{r echo=FALSE, fig.width = 10}
xVector1 <- LogisticLoop(0.2, 3.5, iter = 25)
xMean1 <- mean(xVector1)
xVector2 <- LogisticLoop(0.5, 3.5, iter = 25)
xMean2 <- mean(xVector2)
xVector3 <- LogisticLoop(0.7, 3.5, iter = 25)
xMean3 <- mean(xVector3)

par(mfrow = c(1,3))
plot(xVector1, xlab = "Iterations",ylab = "Population", col = "red", ylim = c(0,1), main="X0 = 0.2")
lines(xVector1, col = "red", lwd = 1)
abline(h = xMean1, col = "blue")

plot(xVector2, xlab = "Iterations",ylab = "Population", col = "red", ylim = c(0,1), main="X0 = 0.5")
lines(xVector2, col = "red", lwd = 1)
abline(h = xMean2, col = "blue")

plot(xVector3, xlab = "Iterations",ylab = "Population", col = "red", ylim = c(0,1), main="X0 = 0.7")
lines(xVector3, col = "red", lwd = 1)
abline(h = xMean3, col = "blue")
```

## Stable Solutions

We know that at r = 3, we begin to chaotic behaviour. But where are the stable solutions for each r? Here they are.

And our ShinyApp can let you interact with this.

```{r echo=FALSE, fig.width = 10}
solutions <- StableSolutions()

stableMeans <- solutions %>% group_by(R) %>% summarise(mean(X))
        
plot(solutions, xlab = "Growth Rate",
    ylab = "Population", col = "red",
    pch = ".",
    main = "Stable Solutions of the Logistic Equation")
lines(stableMeans, col = "blue")
```
